---
params: 
  taxon_id: "2491123" # sagbrush sparrow
  # taxon_id: "2474941" # can't remember
  # taxon_id: "5229498" Pinyon Jay
  # taxon_id: "2439561" # Olive-backed Pocketmouse
  # taxon_id: "9147984" 
  # taxon_id: "1034935"
format:
  docx:
    reference-doc: 'template_styles/template_mpsg_styles.docx'
    df-print: kable
    knitr:
      opts_chunk:
        dev: 'ragg_png'
        fig-width: 6.5
execute:
  echo: false
  message: false
  warning: false
  class-output: 'r'
  fig-width: 6.5
---

```{r}
library(tidyverse)
library(sf)
library(targets)
library(glue)
library(ggtext)
library(ggspatial)
library(extrafont)
loadfonts(quiet = T)

# t_id<- "2491123" # sagbrush sparrow
# t_id<- "2474941" # whooping crane
# t_id<- "5229498" Pinyon Jay
# t_id<- "2439561" # Olive-backed Pocketmouse
# t_id<- "2703686" # Bromus maginatus
# t_id<- "2362460" # Flathead Chub (NHP Data)
# t_id<-"1034935"  # Prairie Long-lipped Tiger Beeter (GBIF Map data)
# t_id<-"9147984" # Indian Creek Tiger Beetle

t_id <- params$taxon_id
species_data <- tar_read(output_dne_eligible_lists) |>
  filter(taxon_id == t_id) |>
  mutate(
    min_overall_year = min(c_across(contains("minYear")), na.rm = T)
  )

sp_names <- glue("{species_data$common_name} (*{species_data$scientific_name}*)")
```
:::{custom-style="Body Text"}
# `r species_data$common_name` (*`r species_data$scientific_name`*)

## Specialist Recommendation:

### MPSG Specialist Recommendation 
[DEFAULTS: Likely to be of concern/Not likely to be of concern/Limited information to inform recommendation] 

### Rationale 

[`r str_to_upper('Short narrative – see Example rationales in Lolo overviews:')` [2023 Lolo SCC animal overviews.pdf](https://usdagcc.sharepoint.com/:b:/r/sites/fs-cio-mpsg/SCC Library/General Resources/Species overviews by region/R1/2023 Lolo SCC animal overviews.pdf?csf=1&web=1&e=IJFJXn), [2023 Lolo SCC plant overviews.pdf](https://usdagcc.sharepoint.com/:b:/r/sites/fs-cio-mpsg/SCC Library/General Resources/Species overviews by region/R1/2023 Lolo SCC plant overviews.pdf?csf=1&web=1&e=bpa1G0), [2023 Lolo SCC plant overviews.pdf](https://usdagcc.sharepoint.com/:b:/r/sites/fs-cio-mpsg/SCC Library/General Resources/Species overviews by region/R1/2023 Lolo SCC plant overviews.pdf?csf=1&web=1&e=bpa1G0), and [Determination Rational Support.xlsx](https://usdagcc.sharepoint.com/:x:/r/sites/fs-cio-mpsg/SCC Library/General Resources/Determination Rational Support.xlsx?d=wdbc0bc6f4b9f476ab31cf3bc0729fe1f&csf=1&web=1&e=5ZlEyW).]


## Regional Summary:

## Species occurrence records and conservation categories

### What, if any, are the scientific name synonyms for the species?

```{r}
synonyms <- tar_read(eligible_synonyms) |>
  filter(taxon_id == t_id) |>
  pull(canonicalName)

citation <- "{GBIF Secretariat 2023}"

if (length(synonyms) > 0) {
  synonyms <- synonyms |>
    paste(collapse = ", ")

  synonyms <- glue("Current and historic synonyms for *{species_data$scientific_name}* are: *{synonyms}* [citation].")
} else {
  synonyms <- glue("There are no known synonyms for this species [citation].")
}
```

`r synonyms`

### NatureServe Conservation Status (Global/State):

`r glue("{species_data$rounded_gRank}/{species_data$NE_sRank}(NE), {species_data$SD_sRank}(SD)")`

### Additional Qualifying Conservation Categories (including Local Concern)

```{r}
other_ranking_narrative <- if (
  !is.na(species_data$r2_ss_list) |
    !is.na(species_data$nebraska_swap) |
    !is.na(species_data$sd_te) |
    !is.na(species_data$usfws_status)
) {
  species_data |>
    mutate(
      other_rankings = glue(
        ifelse(!is.na(r2_ss_list), "Region 2 Sensitive Species, ", ""),
        ifelse(!is.na(nebraska_swap), "Nebraska SWAP {nebraska_swap}, ", ""),
        ifelse(!is.na(sd_te), "South Dakota {sd_te}", ""),
        ifelse(!is.na(usfws_status), "FWS {usfws_status}, ", "")
      ),
      other_rankings = str_replace(other_rankings, ".{2}$", ".")
    ) |>
    pull(other_rankings)
} else {
  "No other qualifying conservation categories apply to this species."
}
```

`r other_ranking_narrative`

### Number of Occurrence Records by Data Source

```{r}
number_of_occ <- species_data |>
  mutate(
    occ_by_source = glue(
      "Occurrence: ",
      ifelse(!is.na(GBIF_nObs) & GBIF_nObs != 0, "GBIF - {GBIF_nObs}, ", ""),
      ifelse(!is.na(iDB_nObs) & iDB_nObs != 0, "iDigBio - {iDB_nObs}, ", ""),
      ifelse(!is.na(SEI_nObs) & SEI_nObs != 0, "SEINet - {SEI_nObs}, ", ""),
      ifelse(!is.na(IMBCR_nObs) & IMBCR_nObs != 0, "Bird Conservancy of the Rockies - {IMBCR_nObs}, ", ""),
      ifelse(!is.na(NENHP_nObs) & NENHP_nObs != 0, "Nebraska Natural Heritage - {NENHP_nObs}, ", ""),
      ifelse(!is.na(SDNHP_nObs) & SDNHP_nObs != 0, "South Dakota Natural Heritage - {SDNHP_nObs}, ", "")
    )
  ) |>
  pull(occ_by_source) |>
  str_replace(".{2}$", ".")
```
`r number_of_occ`

#### Year of first and last observation for all occurrence data: 

First: `r species_data$min_overall_year`, Last: `r species_data$max_overall_year`

### Is the species currently federally designated as threatened, endangered, proposed, or candidate species under the Endangered Species Act?

```{r}
if (is.na(species_data$usfws_status)) {
  fed_status_string <- "No"
} else if (species_data$usfws_status == "Under Review") {
  fed_status_string <- glue("No. Current USFWS status: {species_data$usfws_status}.")
} else {
  fed_status_string <- glue("Yes. Current USFWS status: {species_data$usfws_status}.")
}
```

`r fed_status_string`

### Species' Native Range 

```{r}
## Base plot style.
p_base <- ggplot() +
  theme_minimal(base_family = "Roboto Condensed") +
  ggspatial::annotation_north_arrow(
    location = "br",
    which_north = "true",
    height = unit(0.33, "in"),
    width = unit(0.33, "in"),
    style = north_arrow_fancy_orienteering()
  ) +
  theme(
    plot.background = element_rect(color = "black", linewidth = 1),
    plot.title = element_markdown(face = "bold", hjust = 0, margin = margin(0, 0, 5, 0, "pt")),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, "pt"),
    plot.caption = element_markdown(),
    legend.title = element_markdown(),
    legend.position = "inside",
    legend.position.inside = c(0.15, 0.1),
    legend.box.background = element_rect(color = "black", linewidth = 0.5, fill = "white"),
    legend.box.margin = margin(0, 3, 0, 3, unit = "pt")
  )

base_map_data <- tar_read(evaluation_base_map_data)
```


```{r}    
#| label: cont-range
#| fig-cap: "Species range, clipped to show only the continental United States (note this may not reflect the species’ full range size)."
#| fig-alt: "Species range, clipped to show only the continental United States (note this may not reflect the species’ full range size)."
#| fig-height: 4.85


m_source <- tar_read(map_source) |>
  filter(taxon_id == t_id) |>
  pull(source)

### Need to fix so that GBIF occ data gets pulled. Now there is no NA in map source.
if (is.na(m_source)) {
  m_source <- "check_gbif"
}

if (m_source == "BIEN" | m_source == "IUCN") {
  base_area <- base_map_data$l_48

  aoa <- tar_read(nnfg_aoa) |>
    st_transform(st_crs(base_area)) |>
    st_union() |>
    st_as_sf()

  aoa_center <- aoa |>
    st_centroid() |>
    st_coordinates() |>
    as_tibble()


  if (m_source == "BIEN") {
    g_map_data <- tar_read(bien_plant_maps)
    source_statement <- "**Source**:Botanical Information and Ecology Network."
  }
  if (m_source == "IUCN") {
    g_map_data <- tar_read(all_iucn_map)
    source_statement <- "**Source: IUCN Spatial Data**"
  }

  g_map_data <- g_map_data |>
    filter(taxon_id == t_id) |>
    st_transform(st_crs(base_area)) |>
    st_intersection(st_union(base_area))

  p_base +
    geom_sf(data = base_area, fill = "grey30", color = "white") +
    geom_sf(data = g_map_data, aes(fill = "Species Range"), color = "transparent", alpha = 0.8) +
    geom_sf(data = aoa, fill = "yellow", color = "transparent") +
    annotate(
      geom = "curve",
      x = aoa_center$X + 950000,
      y = aoa_center$Y - 950000,
      xend = aoa_center$X,
      yend = aoa_center$Y,
      color = "#d35400",
      linewidth = 1.1,
      arrow = arrow(length = unit(2, "mm"))
    ) +
    geom_sf_label(
      data = aoa,
      label = "Nebraska National\nForests and Grasslands",
      size = 2.5,
      nudge_x = 950000, nudge_y = -950000,
      fill = "#edbb99"
    ) +
    scale_fill_manual(values = c("#1e8449")) +
    labs(
      title = sp_names,
      subtitle = "Continental United States Range",
      fill = "**Legend**",
      caption = source_statement,
      x = "",
      y = ""
    )
}
```

```{r}
if (m_source == "EBIRD" | m_source == "check_gbif") {
  base_area <- base_map_data$north_america

  aoa <- tar_read(nnfg_aoa) |>
    st_transform(st_crs(base_area)) |>
    st_union()

  aoa_center <- aoa |>
    st_centroid() |>
    st_coordinates() |>
    as_tibble()

  if (m_source == "check_gbif") {
    g_map_data <- tar_read(gbif_occ_data) |>
      filter(taxon_id == t_id) |>
      st_transform(st_crs(base_area)) |>
      st_intersection(base_area |> st_union())
    if (nrow(g_map_data) != 0) {
      p_base_a <- p_base +
        geom_sf(data = base_area, fill = "grey30", color = "white") +
        geom_sf(data = g_map_data, aes(color = "Occurrence Record"), size = 2) +
        geom_sf(data = aoa, fill = "yellow", color = "transparent") +
        annotate(
          geom = "curve",
          x = aoa_center$X + 950000,
          y = aoa_center$Y - 950000,
          xend = aoa_center$X,
          yend = aoa_center$Y,
          color = "#d35400",
          linewidth = 1.1,
          arrow = arrow(length = unit(2, "mm"))
        ) +
        geom_sf_label(
          data = aoa,
          label = "Nebraska National\nForests and Grasslands",
          size = 2.5,
          nudge_x = 950000, nudge_y = -950000,
          fill = "#edbb99"
        ) +
        scale_color_manual(values = c("#1e8449")) +
        labs(
          title = sp_names,
          subtitle = "North American Range",
          color = "**Legend**",
          caption = "**Source:** GBIF occurrence records.",
          x = "",
          y = ""
        )
    } else {
      m_source <- "no_maps"
    }
  } else if (m_source == "EBIRD") {
    g_map_data <- tar_read(bird_maps) |>
      filter(taxon_id == t_id) |>
      st_transform(st_crs(base_area)) |>
      st_intersection(base_area |> st_union())

    count_season <- g_map_data |>
      as_tibble() |>
      count(season) |>
      pull(season) |>
      length()

    if (count_season > 1) {
      mig_orig <- g_map_data |>
        filter(season == "Migration")

      breed_orig <- g_map_data |>
        filter(season == "Breeding")

      nonbreed <- g_map_data |>
        filter(season == "Nonbreeding")

      year_round <- breed_orig |>
        st_intersection(nonbreed) |>
        st_intersection(mig_orig) |>
        mutate(
          season = "Year Round"
        ) |>
        select(season)

      g_map_data <- g_map_data |>
        select(season) |>
        bind_rows(year_round)
    } else {
      g_map_data <- g_map_data |>
        mutate(season = "Year Round")
    }

    p_base_a <- p_base +
      geom_sf(data = base_area, fill = "grey30", color = "white") +
      geom_sf(data = g_map_data, aes(fill = season), color = "transparent") +
      scale_fill_brewer(palette = "Dark2") +
      geom_sf(data = base_area, fill = "transparent", color = "white") +
      labs(
        title = sp_names,
        subtitle = "North American Range",
        fill = "**Season**",
        caption = "**Source:** eBird Status and Trends"
      )
  }
}
```
```{r}
#| label: breeding-range-year-round-gbif
#| fig-cap: "North American Seasonal Ranges"
#| fig-alt: "North American Seasonal Ranges"
#| fig-height: 6.5

if (m_source == "check_gbif") {
  p_base_a
}
```


```{r}
#| label: breeding-range-seasonal
#| fig-cap: "North American Seasonal Ranges"
#| fig-alt: "North American Seasonal Ranges"
#| fig-height: 7
if (m_source == "EBIRD") {
  if (count_season > 1 & nrow(g_map_data) > 0) {
    p_base_a +
      facet_wrap(~season) +
      theme(
        legend.position = "none"
      )
  }
}
```

```{r}
#| label: year-round-range
#| fig-cap: "North American Year Round Ranges"
#| fig-alt: "North American Year Round Ranges"
#| fig-height: 6.5
if (m_source == "EBIRD") {
  if (count_season == 1 & nrow(g_map_data) > 0) {
    p_base_a +
      geom_sf(data = aoa, fill = "yellow", color = "transparent") +
      annotate(
        geom = "curve",
        x = aoa_center$X + 950000,
        y = aoa_center$Y - 950000,
        xend = aoa_center$X,
        yend = aoa_center$Y,
        color = "#d35400",
        linewidth = 1.1,
        arrow = arrow(length = unit(2, "mm"))
      ) +
      geom_sf_label(
        data = aoa,
        label = "Nebraska National\nForests and Grasslands",
        size = 2.5,
        nudge_x = 950000, nudge_y = -950000,
        fill = "#edbb99"
      )
  }
}
```


```{r}
#| label: occurrence-records
#| fig-cap: "Occurrence records within Nebraska National Forests and Grasslands."
#| fig-alt: "North American Seasonal Ranges"
#| fig-height: 6.35

sp_points <- tar_read(all_eligible_spatial_data_point) |>
  filter(taxon_id == t_id)

sp_polys <- tar_read(all_eligible_spatial_data_poly) |>
  filter(taxon_id == t_id)

unit_sp <- tar_read(nnfg_bd)
own <- tar_read(nnfg_fs_ownership)

labels <- base_map_data$nnfg_dist %>%
  bind_cols(st_coordinates(st_centroid(.))) |>
  mutate(
    name_sub = str_replace(DISTRICTNAME, " Ranger District", ""),
    label_n = glue("{name_sub} RD"),
    X = case_when(
      OBJECTID == 429949 ~ 850000,
      OBJECTID == 429950 ~ 630000,
      OBJECTID == 429951 ~ 620000,
      TRUE ~ X
    ),
    Y = case_when(
      OBJECTID == 429949 ~ 4700000,
      name_sub == "Wall" ~ 4885000,
      name_sub == "Pine Ridge" ~ 4700000,
      name_sub == "Ft. Pierre" ~ 4870000,
      name_sub == "Fall River" ~ 4805000,
      TRUE ~ Y
    )
  )

roads <- base_map_data$nnfg_roads |>
  select(highway, `tiger:name_base`:`tiger:name_base_4`)
motorways <- roads |>
  filter(highway == "motorway") |>
  mutate(name = case_when(
    `tiger:name_base` == "I-190" ~ "I-90",
    TRUE ~ `tiger:name_base`
  )) |>
  count(name)

trunk <- roads |>
  filter(highway == "trunk")

other_rd <- roads |>
  filter(highway %in% c("primary"))

states <- base_map_data$l_48 |>
  st_transform(st_crs(unit_sp)) |>
  filter(name %in% c("Nebraska", "South Dakota", "Wyoming"))

state_labels <- tibble(
  label = c("South Dakota", "Nebraska"),
  x = c(885000, 885000),
  y = c(4777000, 4767000)
)
wyoming_label <- tibble(
  label = c("Wyoming"),
  x = c(576517.7),
  y = c(4900000)
)

p_base +
  geom_sf(data = unit_sp, fill = "grey70", color = "transparent") +
  geom_sf(data = own, aes(fill = "FS Ownership"), color = "transparent", alpha = 0.5) +
  geom_sf(data = trunk, color = "grey80", linewidth = 1) +
  geom_sf(data = other_rd, color = "grey80", linewidth = 0.5) +
  geom_sf(data = motorways, color = "grey30", linewidth = 1) +
  geom_sf_label(
    data = motorways,
    aes(label = name),
    fill = "black",
    color = "white",
    nudge_x = -33300,
    nudge_y = 18000,
    label.r = unit(0.8, "lines"),
    label.padding = unit(0.35, "lines")
  ) +
  geom_sf(data = states, color = "black", linewidth = 1, fill = "transparent", linetype = "dotdash") +
  geom_sf(data = labels, fill = "transparent", color = "grey30") +
  geom_label(data = labels, aes(X, Y, label = label_n), show.legend = FALSE, size = 3) +
  geom_text(data = state_labels, aes(x, y, label = label), size = 3, color = "#999999") +
  geom_text(data = wyoming_label, aes(x, y, label = label), size = 3, color = "#999999", angle = 90, nudge_x = -7500) +
  geom_sf(data = sp_polys, fill = "#7d3c98", color = "#7d3c98", alpha = 0.7) +
  geom_sf(data = sp_points, size = 2.5, aes(color = "Occurrence"), alpha = 0.7) +
  scale_color_manual(values = c("#7d3c98")) +
  scale_fill_manual(values = c("darkgreen")) +
  labs(
    title = sp_names,
    subtitle = "Occurrence records within the Nebraska National Forests and Grasslands",
    color = "**Database Occurrences**",
    fill = "**Ownership**",
    x = "",
    y = ""
  ) +
  scale_x_continuous(limits = c(567517.7, 899630)) +
  scale_y_continuous(limits = c(4624402.1, 4914917)) +
  theme(
    axis.text = element_blank(),
    panel.grid = element_blank(),
    legend.key.spacing = unit(0, "pt"),
    legend.margin = margin(0, 0, 0, 0, "pt"),
    legend.box.margin = margin(5, 5, 5, 5, "pt")
  )
```


### Is the species native to, and known to occur in, the plan area?

```{r}
nn_base <- species_data |>
  pull(`Is the Species Native and Known to Occur`)

nn_basi <- species_data |>
  pull(`What is the rationale and supporting BASI for recommending that an observation does not meet the requirements of native to, and known to occur in the plan area?`)
```

`r nn_base`


`r ifelse(is.na(nn_basi), "", nn_basi)`

## Species information

### Species Current Range Size and Configuration

#### What is the size of the species' range?

[SELECT FROM: Large (i.e., roughly the area of the continental U.S. or larger), Moderate (i.e., roughly the ½ the area of the continental U.S.), Regional Endemic (i.e., occupying roughly the area of 2-4 states), State Endemic (i.e., only known to occur in the state overlapping the plan area but occupying an area larger than the environs of the plan area), Local Endemic (i.e., a state endemic only found within the plan area and the surrounding counties)]


#### What is the distance between the plan area and the closest edge of the species range? 

[SELECT FROM: unknown (i.e., no known range for the species), 0 miles (i.e. species range transects the plan area), 1-50 miles, 51-150 miles, 151-500 miles, 500 or more miles.]   

#### Is the portion of the species range that overlaps the plan area disjunct from the main range of the species (i.e., the range overlapping the plan area is smaller and not connected to the larger range of the species)?

[SELECT FROM: Yes, No, Unknown]

### Species’ Landscape-scale Habitat

#### What are the landscape-scale habitat types or ecosystems the species occupies based on Best Available Scientific Information (BASI), including literature from outside of the plan area?

```{r}
ns_habs_all <- tar_read(ns_habitats)

ns_habs <- ns_habs_all |>
  filter(taxon_id == t_id)

comments <- ns_habs |>
  filter(hab_cat == "comments")

if (nrow(comments) == 1) {
  comment_narrative <- comments |>
    pull(ns_hab)
} else {
  comment_narrative <- "[NO HABITAT NARRATIVE WAS GENERATED FROM NATURESERVE]"
}
```

**Nature Serve Habitat Information:** *"`r comment_narrative`"* {Natureserve 2024}

[`r str_to_upper("Narrative description pulled in from NatureServe – will need manual review to ensure fine-scale habitat features are not mixed in with landscape, and additional BASI as appropriate; NatureServe also doesn’t capture many plants, so this may also need to be written manually.")`]

#### Based on BASI about the landscape-scale habitat types or ecosystems the species occupies, what landscape-scale habitat types or ecosystems in the plan area are likely to support the species?

```{r}
get_hab_types_all <- tar_read(crosswalk_habitats_to_species)

all_hab_types_sentence <- get_hab_types_all |>
  distinct(mpsg_habitat_bin) |>
  filter(!is.na(mpsg_habitat_bin)) |>
  pull(mpsg_habitat_bin) |>
  paste(collapse = "; ")

get_hab_types <- get_hab_types_all |>
  filter(taxon_id == t_id) |>
  filter(!is.na(mpsg_habitat_bin))

if (nrow(get_hab_types) > 0) {
  hab_type_narrative <- get_hab_types |>
    distinct(mpsg_habitat_bin) |>
    pull(mpsg_habitat_bin) |>
    paste(collapse = ", ")
} else {
  no_data_warning <- str_to_upper("There are no habitat types available for this species from Naturserve that crosswalk to MPSG defined ecosystems.")
  hab_type_narrative <- glue("{no_data_warning} THE FOLLOWING ARE MPSG DEFINED ECOSYSTEMS: {all_hab_types_sentence}.")
}
```

`r hab_type_narrative`


#### Does the species respond uniquely to landscape-scale habitat threats and trends, and if so, how? 

`r str_to_upper("Manually review and bring in any species-specific info about how the species responds to general habitat trends/threats. The appendix explains the macrohabitat availability, distribution, and threats, but this is where you would connect it to the species- e.g., if fire is a threat, but species actually needs fire for its seeds to germinate, or it actually uses roadside habitat so road development is not as much of a threat")`].

The general trends and threats of this species’ landscape-scale habitats are discussed in the Appendix, and there is no evidence that this species responds in any unique ways to the landscape-scale habitat threats and trends.

### Species Fine-scale Habitat

#### What are the fine-scale habitat features or conditions that the species relies on? 

[`r str_to_upper("Narrative description from literature. To help focus the evaluation on issues that are likely to affect a species long-term persistence the inclusion of fine-scale habitat conditions or features here, should be limited to consideration of conditions or features that: 1) BASI indicates are limited in abundance or distribution, 2) BASI suggest are declining in abundance or distribution, or 3) BASI indicates have a disproportionate value to the population dynamics and distribution of the species because of the species specialized association with the feature or life history dependence (e.g., rare nesting substrate, rare soil type, cold water refugia).")`]

#### What are the threats and trends for the fine-scale habitat features this species relies on (distribution, abundance, and quality), both within and beyond the plan area?
[DEFAULT: There are no known specific population trends for the species in the plan area. ]

[`r str_to_upper("Beyond threats documented across the species range {CITATIONS}, there are no known unique threats to the species within the plan area.")`]

### Population 

#### Abundance and distribution of species in the plan area

[DEFAULT: No information could be found regarding the distribution, abundance and quality of this species’ microhabitat in the plan area.]

#### Population trend of species on and off the plan area

```{r}
if (!is.na(species_data$class) & species_data$class == "Aves") {
  imbcr_narrative <- tar_read(imbcr_trend_narratives)$unit_narratives |>
    filter(species == species_data$common_name)

  if (nrow(imbcr_narrative) == 1) {
    imbcr_narrative <- imbcr_narrative |>
      pull(narrative)
  } else {
    imbcr_narrative <- "No IMBCR trend information available for this species."
  }

  imbcr_additional_narrative <- tar_read(imbcr_trend_narratives)$additional_narratives |>
    filter(species == species_data$common_name)

  if (nrow(imbcr_additional_narrative) == 1) {
    imbcr_additional_narrative <- imbcr_additional_narrative |>
      pull(narrative)
  } else {
    imbcr_additional_narrative <- NA
  }

  bbs_narrative <- tar_read(bbs_trend_narratives) |>
    filter(species_name == species_data$common_name)

  if (nrow(bbs_narrative) == 1) {
    bbs_narrative <- bbs_narrative |>
      pull(final_narrative)
  } else {
    bbs_narrative <- "No narrative"
  }
}
```

`r if(!is.na(species_data$class) &species_data$class=="Aves"){ ifelse(imbcr_narrative!="No narrative", imbcr_narrative, "")}`

`r if(!is.na(species_data$class) &species_data$class=="Aves"){ ifelse(!is.na(imbcr_additional_narrative), glue("**Additional Narratives you can add if you want or if there are no narritives above**: {imbcr_additional_narrative}"), "")}`

`r if(!is.na(species_data$class) &species_data$class=="Aves"){ ifelse(bbs_narrative!="No narrative", bbs_narrative, "")}`

[DEFAULT: There are no known population estimates for the species in the plan area. Beyond threats documented across the species range [CITATIONS], there are no known unique threats to the species within the plan area.]

#### Non-habitat-mediated threats to populations (on or off the plan area)

[`r str_to_upper("Consider disease, competition, contamination, invasives")`]



### Natural and Life History Characteristics that Inform Species Persistence 

#### Dispersal ability

[`r str_to_upper("Brief, ~ one sentence description of species’ movement scale and its overall dispersal capacity")`]

#### Life history strategy

[`r str_to_upper("Brief, ~ one sentence description of species’ life history, recognizing species generally fit one of these categories (equivalent is annual/perennial for plants): Slow: large body size, long generation, high survival, low reproductive investment; Fast: small body size, short generation, low survival, high reproductive investment")`] 

#### Ecological specialization 

[`r str_to_upper('Habitat or Substrate Specialist, Dietary Specialist, Symbiotic (mutualistic, parasitic, or commensal)')`]

## References

:::



